---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dynamic

<!-- badges: start -->
<!-- badges: end -->

The goal of dynamic is to simulate fit index cutoffs for latent variable models that are tailored to the user's model statement, model type, and sample size. 

This is the counterpart of the Shiny Application, [dynamicfit.app](https://dynamicfit.app/cfa).  The Shiny app and the R package will give you the same results.  If you are comfortable with R, consider using the package during high traffic times to reduce server burden.

## Installation

This is the beta version of the package.  Please submit bug reports and issues on GitHub.  You can install the beta version of dynamic from [Github](https://github.com) with:

```r
library(devtools)
devtools::install_github("melissagwolf/dynamic")
```

```{r echo=FALSE,eval=FALSE}
#Add this when it's on CRAN

#You can install the released version of dynamic from #[CRAN](https://CRAN.R-project.org) with:

#install.packages("dynamic")
```

## Example

```{r example, eval=FALSE}
library(dynamic)
mod <- "F1 =~ .602*Y1 + .805*Y2 + .516*Y3 + .857*Y4
        F2 =~ .413*Y5 + -.631*Y6
        F1 ~~ .443*F2
        Y4 ~~ .301*Y5"
cfaFit(mod,500)
```
